Functional Analyses of Distinct Oxygen Concentrations and Light Niches of *Prochlorococcus marinus*

### 2024-02-14 Code/Manuscript_O2.Rmd
Edits to Materials & Methods

## Culturing {.unnumbered}

Three xenic *P. marinus* cultures, obtained from Bigelow Labs, NCMA
Maine, US: MED4 (CCMP1986) from a High Light-adapted (HLI) clade, SS120
(CCMP1375) from a Low Light-adapted (LLIII) clade and MIT9313 (CCMP2773)
also from a Low Light-adapted (LLIV) clade. These cultures were then
maintained in two separate incubators. The temperature for both
incubators was set to 22°C and a light/dark cycle of 12 h. The PAR level
of the incubator was chosen to reflect the light level of the natural
niche of the ecotype during culturation. The PAR level of the incubator
containing the HLI clade, MED4, was set at 160 µmol photons m^-2^ s^-1^,
whereas the incubator containing the LLIII and LLIV clades, SS120 and
MIT9313 respectively, was set at 30 µmol photons m^-2^ s^-1^. To ensure
cultures remained in exponential growth phase, all strains were
transferred weekly in Pro99 media prepared according to
[@mooreCulturingMarineCyanobacterium2007] in autoclaved artificial
seawater. Artificial seawater was prepared according to the National
Center for Marine Algae and Microbiota (NCMA) protocol by combining salt
solution I and salt solution II using the enriched artificial seawater
(ESAW) recipe.

## Experimental design {.unnumbered}

For each of three ecotypes, we imposed [O~2~] (three levels: 2.5 µM,
25µM, 250 µM), photoperiod (four levels: 4 h, 8 h, 12 h, 16 h), spectral
waveband (four bands: full spectrum white LED, 660 nm, 530 nm, 450 nm),
and light level (three levels: 30, 90, 180 µmol photons m^-2^ s^-1^)
treatments in a factorial design. Each factor is explained below. The
full crossing of all factors would yield 3 x 4 x 4 x 3 = 144 treatments
per ecotype (432 total), but due to time constraints and total absence
of growth of some ecotypes under some conditions, not all treatments
were carried out. In total, we completed 291 treatments.

All growth experiments were conducted at 22°C.

## Experimental light conditions {.unnumbered}

Three growth Photosynthetically Active Radiation (PAR) levels (180, 90,
and 30 µmol photons m^-2^ s^-1^) and four spectral wavebands (white LED
full spectrum, 660 nm, 530 nm, and 450 nm) were chosen to simulate light
levels and spectral color spanning the vertical ocean water column, from
near-surface to the lower euphotic zone depths. For simplicity, actinic light used for growth
under specific wavebands will be referred by the respective spectral
color; white LED for LED full spectrum, red for 660 nm, green for 530 nm
and blue for 450 nm. Four different photoperiods were chosen to simulate
various diel cycles characteristic of current and hypothetical future
niches of *P. marinus*. A photoperiod of 16 h was chosen to represent
temperate (45°N) summer at the ocean surface, 12 h for equatorial (0°N)
ocean surface or temperate (45°N) spring and fall ocean surface or
temperate (45°N) summer at deep ocean depths, 8 h for temperate (45°N)
winter at the surface or at temperate (45°N) spring and fall at depth
and equatorial (0°N) deep ocean depths and 4 h for temperate (45°N)
winter or deep ocean depths during temperate (45°N) spring and fall.

## Experimental oxygen conditions {.unnumbered}

Three target dissolved oxygen concentrations [O~2~] were delivered to
tubes of the Multicultivator by mixing varying ratios of air and
Nitrogen (N~2~) gases while delivering 0.05% of Carbon Dioxide (CO~2~)
gas through a 0.2 μm sterile microfilter via a G400 gas mixing system.
To confirm and monitor the [O~2~], 4 FireSting optodes (PyroScience,
Germany) were inserted into select tubes of each modified [O~2~] run for
real-time measurements. A compensation temperature probe was placed in
the aquarium of the bioreactor to correct [O~2~] for temperature
fluctuations. In addition, the software corrected [O~2~] based on the
salinity of the media (32 ppt). For the low O~2~ environment
experiments, 0.5 to 1.0 µM of O~2~ was delivered to each Multicultivator
tube by sparging with a gas mixture containing 99.95% N~2~ and 0.05%
CO~2~ to purge dissolved O~2~ out of the culture. The intermediate O~2~
environment experiments were sparged to deliver 10 to 25 µM of O~2~
using a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. The
high O~2~ environment experiments were sparged with lab air (78% N~2~,
21% O~2~, 1% Ar and 0.05% CO~2~) to deliver 230 µM of O~2~. While the
flow rate of the gas mixture was controlled, variations in bubbling
speed affected the [O~2~] delivered to each tube; therefore, a range of
[O~2~] was defined for each experimental O~2~ level; 0.5 µM - 5 µM for
low, 5 µM - 50 µM for intermediate and 200 µM - 280 µM for high O~2~
experiments. To simplify the representation of experimental [O~2~]
conditions in graphs and discussions, the approximate median [O~2~] of
each experimental range: 2.5 µM, 25 µM, and 250 µM, will be used for
low, intermediate, and high conditions, respectively.

## Bioreactors {.unnumbered}

Growth experiments under different spectral wavebands were performed
using a PSI MCMIX-OD Multicultivator (Figure \@ref(fig:LabeledMC)) or a
PSI MC1000-OD Multicultivator for white LED experiments. Each
Multicultivator has the capacity to individually control 8 tubes with
specific PAR levels and photoperiods and the MCMIX-OD has options for
individually controlled spectral wavebands. 10 mL of exponential growth
culture was added to 70 mL of Pro99 media and all 8 tubes were situated
in a common temperature-controlled water bath to ensure the temperature
remained constant at 22°C over the duration of the experiment. Real time
absorbance measurements of Optical Density (OD) 680 nm (a proxy for cell
suspension density, cell scatter and cell chlorophyll content) and OD
720 nm (a proxy for cell suspension density and cell scatter) were
recorded every 5 minutes for at least 8 to 14 days depending on the
duration of the lag phase, if any. Real time OD
measurements eliminate intrusive subsampling of sterile cultures and
provide high resolution chlorophyll and cell scatter proxies over the
duration of the experiment. All data from the Multicultivator were saved
as a comma separated values file and processed in R-Studio for
calculations of growth rate estimates and graphical plotting.

>>>>

## Ocean Protein Portal analyses {.unnumbered}

## *Prochlorococcus* culturing {.unnumbered}

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA
Maine, US (REFERENCE). MED4 (CCMP1986) is from High Light-adapted (HLI) clade; SS120
(CCMP1375) is from Low Light-adapted (LLIII) clade; and MIT9313 (CCMP2773)
is from Low Light-adapted (LLIV) clade. Cultures were
maintained in incubators set to 22°C with a light/dark cycle of 12 h, with illumination from XXXXX. The PAR level for maintenance cultures reflected  Photosynthetically Active Radiation (PAR) in the source
niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^; SS120 and
MIT9313 at 30 µmol photons m^-2^ s^-1^. To maintain active growth all strains were
transferred weekly with XX to YY dilution with Pro99 media
[@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial
seawater (National
Center for Marine Algae and Microbiota (NCMA) protocol by combining salt
solution I and salt solution II using the enriched artificial seawater
(ESAW) recipe.  REPLACe with CITATION OF NCMA protocol)XXXX

## Experimental design using bioreactor control of light and [O~2~] {.unnumbered}

Controlled growth experiments were performed using PSI Multicultivators (Figure \@ref(fig:LabeledMC); MCMIX-OD orMC1000-OD, Brno, Czech Republic ).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (AA, BB, CC, DD nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~)
gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system XXXXGIVE SOURCEXXXX. [O~2~] in situ was verified using oxygen optodes (PyroScience,
Germany) inserted into tubes for
real-time measurements, with a temperature probe in
the aquarium of the bioreactor to correct [O~2~] measures for temperature
fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the
salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05%
CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

The full crossing of all factor levels would yield 4 x 3 x 4 x 3 = 144 treatments, x 3 strains for 432 possible combinations. Consistent absence of growth of some strains under some levels of photoperiod, PAR, or [O~2~] meant we completed 291 growth factor treatment combinations.

In situ XXXXItalicsXXX measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data from the Multicultivator were saved
as comma separated values file  (CITATION of FILES LOCATION).

PAR of 180, 90 or 30 µmol photons m^-2^ s^-1^, and spectral wavebands (white LED
full spectrum, 660 nm, 530 nm, and 450 nm) were chosen to approximate light
levels and spectral colors spanning the vertical ocean water column, from
near-surface to the lower euphotic zone depths. Photoperiods were chosen to approximate diel cycles characteristic of current and hypothetical future
niches of *P. marinus*; 16 h represents
temperate (45°N) summer at the ocean surface; 12 h for equatorial (0°N)
ocean surface or temperate (45°N) spring and fall ocean surface or
temperate (45°N) summer at deeper ocean depths; 8 h for temperate (45°N)
winter at the surface or at temperate (45°N) spring and fall at depth
and equatorial (0°N) deep ocean depths; and 4 h for temperate (45°N)
winter or deep ocean depths during temperate (45°N) spring and fall.
